---
title: "Allometry_testing"
author: "Kelly Heilman"
date: "9/4/2020"
output: pdf_document
---


### Notes about setup--there are two options:
1. Clone the pecan repo directly, then open this rmd in the repository. (Maybe best option because of dependancies)
2. You can install the allometry module from github with the code below (from allometry vignette):
library(devtools)
# # if not run, need to install the pecan modules
install_github("PecanProject/pecan",subdir="base/logger")
install_github("PecanProject/pecan",subdir="allometry")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(PEcAn.allometry)
```



## Basic Pecan allometry tutorial
Alot of this text and code is just annotated from the allometry vignette.
To view the different components of the tree for which you could develop allomtrys + their abbreviations:

```{r cars}

data("allom.components")
allom.components

pfts = list(PIPO = data.frame(spcd=122,acronym='PIPO')) # list our "Pfts--plant functional types" of interest--really list the species

```

## AllomAve function in PEcAn allometry module
- Fits the bayesian allomtric models for the pfts. ngibbs indicates the number of gibbs/mcmc samples
- looks like there is a burn in of ~200
- automatically outputs a pdf with the traceplots titled "allom.pipo.6.mcmc" and saves and .Rdata object.
- not totally clear to be what the different betas are


```{r,  warning=FALSE}
allom.stats = AllomAve(pfts,components =6, ngibbs=1000) 
allom.stats$PIPO
```

## Predict for individual trees:
- the allom.predict fuction will use the fit relationships from allomAve to predict the biomass component for the pft of interest (PIPO) over the DBH vector values:
- dbh = diameter to predict at
- component is which component of biomass to predict see defintation in the allom.components object
- unclear to me whether component = 6 is predicting Stem biomass, or is only suing the diameter of the stem in the allometry equation
- use = Bg 


```{r}

##Predict for individual trees:
allom.fit = load.allom(getwd()) # get the allometry listed in our working directory
dbh = 1:50 # vector of DBH values to predict over

pred = allom.predict(allom.fit, dbh = dbh, pft = "PIPO", component =6, use = "Bg", interval = "prediction") # predict allometries for trees with dbh 1:50, and get the prediction interval
conf = allom.predict(allom.fit,dbh = dbh,pft = "PIPO",component = 6,use = "Bg",interval = "confidence")# predict allometries for trees with dbh 1:50, and get the confidence interval (including uncertainties)
PI = apply(pred,2,quantile,c(0.025,0.5,0.975),na.rm=TRUE)
CI = apply(conf,2,quantile,c(0.025,0.5,0.975),na.rm=TRUE)
plot(dbh,CI[2,],type='l',lwd=3,ylim=range(PI),ylab="Biomass (kg)") +
lines(dbh,CI[1,],lty=2,col="blue") +
lines(dbh,CI[3,],lty=2,col="blue") +
lines(dbh,PI[1,],lty=3,col="red") +
lines(dbh,PI[3,],lty=3,col="red")

```

## other stuff I learned from digging:
- there is a file in the pecan repository that has a table with all the beta parameters from jenkins

- really we want to use the query.allom.data function to query the "field" data, but there is a database that this function connects to via "con=" argument...and I can't figure out where this database is located/what the connection should be
```{r}

jenkins.table <- read.csv(system.file("data/Table3_GTR-NE-319.v2.csv", package = "PEcAn.allometry"), header = TRUE)
#colnames(jenkins.table)[1] <- c("")
head(jenkins.table)

# other functions
#??AllomAve


#??AllomUnitCoef # converts length units FROM cm TO specified units converts mass units TO kg FROM specificed units



#??allom.BayesFit #Module to fit a common power-law allometric model to a mixture of raw data and allometric equations in a Heirarchical Bayes framework with multiple imputation of the allometric data


#??allom.predict # Function for making tree-level Monte Carlo predictions from allometric equations estimated from the PEcAn allometry module



#??load.allom() # load .allom files

# This is proabably what we want
#??query.allom.data# Module to grab allometric information from the raw data table Will grab both original field data and tallied equations
                    #Tallied equation format based on Jenkins et al 2004 USFS General Technical Report NE-319

#??read.allom.data # Extracts PFT- and component-specific data and allometeric equations from the specified files.

# note that query.allom.data is not an exported fuction from the stand alone module....
#query.allom.data("PIPO")
```
